---
title: "Single Mark-Release Experiment, Part III"
author: ""
date: ""
output: html_document
---

```{r, echo=FALSE}
library(data.table)
```
**Section 1.**  
You are asked by a fishery manager to determine sample size recommendations for a mark-recapture experiment. The population is thought to be between 10,000 and 40,000 fish. The manager wants the estimate to be within 25% of the true population size with 95% confidence.

**a.** Find combinations of $n_1$ and $n_2$ for the two extremes of abundance that achieve this goal and produce a nice
smooth curve ($n_2$ versus $n_1$). Also calculate the value for which the two sample sizes are equal.

**Sample size**

Define parameters
```{r}
low.N <- 10000
low.n1 <- 1:low.N
high.N <- 40000
high.n1 <- 1:high.N
D <- 69.9
```
Build a function to estimate
```{r}
RR <- function(D, N.est, n1) {
    N.est/((n1 * (N.est - 1))/((N.est - n1) * D) + 1)
}
```

Calculate $n_1 = N_2$

```{r}
equal.sample <- function(D, N.est) {
    N.est/(sqrt((N.est - 1)/D) + 1)
}

low.e <- equal.sample(D, low.N)
high.e <- equal.sample(D, high.N)

low.e
high.e
```

Plot the estimates

```{r}
plot(low.n1, RR(D, low.N, low.n1), type = "l", lwd = 2, xlim = c(100, 5000), 
    ylim = c(100, 5000), xlab = "n1", ylab = "n2 sample size")
lines(high.n1, RR(D, high.N, high.n1), col = 4, lwd = 2)
legend("topright", inset = 0.05, c("N = 10,000", "N = 40,000"), lty = c(1, 1), 
    col = c(1, 4))
points(low.e, low.e, cex = 3, pch = 19)
points(high.e, high.e, cex = 3, col = 4, pch = 19)
```

**Section 2.**   
This exercise is an example of Monte Carlo simulation. We will evaluate whether a simulation of a single-release, single-recapture experiment is working properly. We will do this by examining means and variances of various random variables.

Input values
```{r}
N <- 2000
n1 <- 200
n2 <- 100
Em2 <- n1 * n2/N
```

Take random draws of $m_2$ from the hypergeometric distribution
```{r}
set.seed(113)  # for reproducibility
m2 <- rhyper(100, n1, N - n1, n2)
```

Chapman's estimate of $N^*$ and $v^*$

```{r}
N.star <- ((n1 + 1) * (n2 + 1)/(m2 + 1)) - 1
v.star <- (n1 + 1) * (n2 + 1) * (n1 - m2) * (n2 - m2)/(((m2 + 1)^2) * (m2 + 
    2))
```

Calculate average, variance, sd, and se of m2, N, v, t-test and CV across replicates:

```{r}
x.m2 <- mean(m2)
var.m2 <- var(m2)
sd.m2 <- sd(m2)
se.m2 <- sd.m2/sqrt(length(m2))
t.m2 <- (x.m2 - Em2)/se.m2
crit.m2 <- qt((1 - 0.05/2), length(m2) - 1)  #divide alpha by 2 to get a two-sided test
p.m2 <- 2 * pt(-abs(t.m2), length(m2) - 1)
cv.m2 <- sd.m2/x.m2

x.N.star <- mean(N.star)
var.N.star <- var(N.star)
sd.N.star <- sd(N.star)
se.N.star <- sd.N.star/sqrt(length(m2))
t.N.star <- (x.N.star - N)/se.N.star
crit.N.star <- qt((1 - 0.05/2), length(N.star) - 1)  #divide alpha by 2 to get a two-sided test
p.N.star <- 2 * pt(-abs(t.N.star), length(N.star) - 1)
cv.N.star <- sd.N.star/x.N.star

x.v.star <- mean(v.star)
var.v.star <- var(v.star)
sd.v.star <- sd(v.star)
se.v.star <- sd.v.star/sqrt(length(m2))
t.v.star <- (x.v.star - var.N.star)/se.v.star
crit.v.star <- qt((1 - 0.05/2), length(v.star) - 1)  #divide alpha by 2 to get a two-sided test
p.v.star <- 2 * pt(-abs(t.v.star), length(v.star) - 1)
cv.v.star <- sd.v.star/x.v.star
```
Results
```{r}
(t1 = data.table(parameter = c("mean", "var", "sd", "se", "t", "t-crit", "P-value", 
    "CV"), m2 = format(c(x.m2, var.m2, sd.m2, se.m2, t.m2, crit.m2, p.m2, cv.m2), 
    scientific = FALSE, digits = 3), `N*` = format(c(x.N.star, var.N.star, sd.N.star, 
    se.N.star, t.N.star, crit.N.star, p.N.star, cv.N.star), scientific = FALSE, 
    digits = 3), `v*` = format(c(x.v.star, var.v.star, sd.v.star, se.v.star, 
    t.v.star, crit.v.star, p.v.star, cv.v.star), scientific = FALSE, digits = 3)))
```

Approximate variance formula

```{r}
(appr.v.N.star <- N^2 * (1/Em2 + 2/Em2^2 + 6/Em2^3))
```

Histograms

```{r}
par(mfrow = c(2, 1))
hist(m2, breaks = 20, col = 4)
hist(N.star, breaks = 20, col = 8)
```

```{r}
sessionInfo()
```


This is a lab exercise from Dr. Terry Quinn  
at the University of Alaska Fairbanks  
and has been worked up in R by:  
Ben Williams  
bcwilliams2@alaska.edu  
ben.williams@alaska.gov


