---
title: "Schnabel Estimation"
output:
  html_document:
    self_contained: no
---
```{r set-options, echo=FALSE, cache=FALSE}
options(width = 95)
library(data.table)
```

Input data and create cfrog dataframe
```{r input}
n <- c(32, 54, 37, 60, 41)  # number of captures
m <- c(0, 18, 31, 47, 36)  # number of recaptures
R <- c(32, 36, 6, 13, 5)  # of marked fish returned to the population
M <- c(0, cumsum(R)[-5])
(cfrog <- data.frame(n = n, m = m, R = R, M = M))
```

Calculate a Petersen estimate *N^* and add it to the cfrog data
```{r petersen}
cfrog$N.hat <- M * n/m  
```

Calculate N* and add it to cfrog data
```{r chapman}
N.star <- ((M + 1) * (n + 1)/(m + 1)) - 1  # Chapman corrected Petersen estimate
cfrog$N.star <- c(NaN, N.star[-1])  # Need to remove the first value as it is an artifact of the Chapman correction
v.star <- (M + 1) * (n + 1) * (M - m) * (n - m)/(((m + 1)^2) * (m + 2))
cfrog$v.star <- c(NaN, v.star[-1])  # Need to remove the first value as it is an artifact of the Chapman correction
cfrog
```

*Mean Petersen Estimates*
```{r mean petersen}
x.N.star <- mean(cfrog$N.star, na.rm = T)
x.v.star1 <- sum(cfrog$v.star, na.rm = T)/(length(cfrog$v.star) - 1)^2
x.v.star2 <- var(cfrog$N.star, na.rm = T)/(length(cfrog$N.star) - 1)
t <- qt((1 - 0.05/2), length(n) - 2)
se.1 <- sqrt(x.v.star1)
se.2 <- sqrt(x.v.star2)
ci.1 <- c(x.N.star - t * se.1, x.N.star + t * se.1)
ci.2 <- c(x.N.star - t * se.2, x.N.star + t * se.2)
```

*Mean Petersen Results*

```{r mpetersen results}
(mP.table = data.table(parameter = c("mean", "var1", "var2", "t", "se1", "se2", 
    "ci.1 low", "ci.1  up", "ci.2 low", "ci.2  up"), value = format(c(x.N.star, 
    x.v.star1, x.v.star2, t, se.1, se.2, ci.1[1], ci.1[2], ci.2[1], ci.2[2]), 
    scientific = FALSE, digits = 2)))
```
**Schnabel Estimate**
```{r schnabel}
nM <- n * M
m.s <- sum(m)
lambda <- sum(nM)
N.schnabel <- lambda/(m.s + 1)
z <- 1.96
ci.sch <- c(lambda * (2 * m.s + z^2 - z * sqrt(4 * m.s + z^2))/(2 * m.s^2), 
    lambda * (2 * m.s + z^2 + z * sqrt(4 * m.s + z^2))/(2 * m.s^2))
```
*Schnabel Results*
```{r sresults}
(mS.table = data.table(parameter = c("N''", "ci low", "ci  up"), value = format(c(N.schnabel, 
    ci.sch[1], ci.sch[2]), scientific = FALSE, digits = 3)))

```
**Schumacher-Eschmeyer**
```{r schumacher-eschmeyer}
y <- m/n
w <- n[-1]
mM <- m[-1] * M[-1]
wM2 <- w * M[-1]^2
s.mM <- sum(mM)
s.wM2 <- sum(wM2)
beta <- s.mM/s.wM2
pred.y <- beta * M
diff <- y - pred.y
wy2 <- w * diff[-1]^2
s.wy2 <- sum(wy2)
N.schesch <- 1/beta
sigsq <- s.wy2/(length(n) - 2)
se.beta <- sqrt(sigsq/s.wM2)
se.N.schesch <- sqrt(1/beta^4 * se.beta^2)
ci.beta <- c(beta - se.beta * t, beta + se.beta * t)
ci.N.schesch <- c(1/ci.beta[2], 1/ci.beta[1])  #invert the limits
```
**Schumacher-Eschmeyer Results**
```{r s-e results}
mSE.table = data.frame(y, c(NA, w), c(NA, mM), c(NA, wM2), pred.y, diff, c(NA, 
    wy2))  # I used a data.frame here instead of a data.table note that they create complementary objects that can be treated in a similar manner
names(mSE.table) <- c("y=m/n", "w", "mM", "wM^2", "Pres y~", "y-y~", "w(y-y~)^2")
mSE.table

mSE.table1 <- data.frame(values = c(beta, N.schesch, sigsq), se = c(se.beta, 
    se.N.schesch, NA), low.CI = c(ci.beta[1], ci.N.schesch[1], NA), up.CI = c(ci.beta[2], 
    ci.N.schesch[2], NA), row.names = c("B~", "N~", "sigsq"))
print(mSE.table1, digits = 2)
```

**MLE Estimate**

*create a function so that it can be optimized (use the absolute value, then can use the standard optimize function that minimizes)*
```{r}
opt.func <- function(N) {
    abs((1 - sum(cfrog$R)/N) - (prod(1 - cfrog$n/N)))
}

optimize(opt.func, c(50, 150))  #optimize over a reasonable range for the population


plot(50:150, abs((1 - sum(cfrog$R)/50:150) - (prod(1 - cfrog$n/50:150))), xlab = "N estimate", 
    ylab = "difference", ylim = c(-0.5, 0.9))  #plot it just for fun
```


**MLE with non-linear search procedure in detail**

*Note that iterations in R are generally poor programming, they are done here simple as an example*
```{r}
fn <- function(x) {
    N <- 97  #starting value of N
    mle = matrix(0, nrow = x, ncol = 8)
    colnames(mle) <- c("N", "g.N", "h.N", "N1", "g.N1", "h.N1", "delta.h", "corr")
    for (i in 1:x) {
        g.N <- prod(1 - n/N)
        h.N <- N - sum(R) - N * g.N
        N1 <- N - 1
        g.N1 <- prod(1 - n/N1)
        h.N1 <- N1 - sum(R) - N * g.N1
        delta.h <- h.N - h.N1
        corr <- -h.N/delta.h
        mle[i, ] <- c(N, g.N, h.N, N1, g.N1, h.N1, delta.h, corr)
        N <- N + corr
    }
    return(mle)
}
fn(5)
```

Results from MLE Estimate

```{r}
N.hat <- fn(5)[5]
N.hat1 <- fn(5)[5] - 1
N.hat2 <- fn(5)[5] - 2
g.Nhat <- prod(1 - n/N.hat)
g.Nhat1 <- prod(1 - n/N.hat1)
g.Nhat2 <- prod(1 - n/N.hat2)
h.Nhat <- N.hat - sum(R) - N.hat * g.Nhat
h.Nhat1 <- N.hat1 - sum(R) - N.hat1 * g.Nhat1
h.Nhat2 <- N.hat2 - sum(R) - N.hat2 * g.Nhat2
D.h <- h.Nhat - h.Nhat1
D.h1 <- h.Nhat1 - h.Nhat2
D2.h <- D.h - D.h1
b.hat <- -0.5 * (N.hat - sum(R)) * D2.h/D.h^2
Nb.hat <- N.hat - b.hat
se <- sqrt((N.hat - sum(R))/D.h)
Nhat.lci <- Nb.hat - 1.96 * se
Nhat.uci <- Nb.hat + 1.96 * se

mle.table <- data.frame(Abundance = c(N.hat, N.hat1, N.hat2), g.N = c(g.Nhat, 
    g.Nhat1, g.Nhat2), h = c(h.Nhat, h.Nhat1, h.Nhat2), Del.h = c(D.h, D.h1, 
    NA), Del2.h = c(D2.h, NA, NA), b.hat = c(b.hat, NA, NA), N.bhat = c(Nb.hat, 
    NA, NA), se = c(se, NA, NA), row.names = c("N^", "N^-1", "N^-2"))
print(mle.table, digits = 2)
```

Now let's do this with a “shorter” method…
There are a number of packages available for these sort of population calculations
You will likely need to download the package before installing them (install.packages())
both fishmethods and the FSA package (http://www.rforge.net/FSA/files/) have functions for calculating the Schabel and Schumacher-Eschmeyer population estimates.

```{r}
library(fishmethods)
schnabel(catch = n, recaps = m, newmarks = R)
```


```{r}
sessionInfo()
```


This is a lab exercise from Dr. Terry Quinn  
at the University of Alaska Fairbanks  
and has been worked up in R by:  
Ben Williams  
bcwilliams2@alaska.edu  
ben.williams@alaska.gov

