---
title: "Lab 1. Single Mark-Release Experiment: Understanding Random Variables"
author: "Ben Williams"
date: ""
output: html_document
---

From $E(m2)={n_{1}n_{2}/N}$ , solve for true abundance $N$. $N$=_____________. Calculate $N$ in the appropriate spot.
Calculate $N^*$ for each replication. 

```{r,echo=FALSE}
library(data.table); library(ggplot2); library(reshape2)
theme_set(theme_bw(base_size=12)+ 
             theme(panel.grid.major = element_blank(), 
                   panel.grid.minor = element_blank()))
```
Starting values and calculate $N$
```{r}
n1 = 2000
n2 = 1000
Em2 = 20
N = n1 * n2/Em2
```
60 Random draws of $m2$
```{r}
set.seed(26)  #so that the values can be reproduced
m2 <- round(rnorm(60, 20, 5))
x.m2 <- mean(m2)
x.m2

sd.m2 <- sd(m2)
se.m2 <- sd.m2/sqrt(length(m2))
lci.m2 <- x.m2 - 1.96 * se.m2
uci.m2 <- x.m2 + 1.96 * se.m2
diff.m2 <- x.m2 - Em2
t.m2 <- diff.m2/se.m2
```

Calculate $N^*$ and associated estimates
```{r}
N_star <- ((n1 + 1) * (n2 + 1)/(m2 + 1)) - 1
x.N_star <- mean(N_star)
sd.N_star <- sd(N_star)
se.N_star <- sd.N_star/sqrt(length(N_star))
lci.N_star <- x.N_star - 1.96 * se.N_star
uci.N_star <- x.N_star + 1.96 * se.N_star
diff.N_star <- x.N_star - N
t.N_star <- diff.N_star/se.N_star
```
$m2$ results
```{r}
m2.table = data.table(parameter = c("mean", "sd", "se", "uci", "lci", "diff", 
    "t-stat"), value = format(c(x.m2, sd.m2, se.m2, lci.m2, uci.m2, diff.m2, 
    t.m2), scientific = FALSE, digits = 2))
m2.table
```

$N^*$ results
```{r}
N.table = data.table(parameter = c("mean", "sd", "se", "uci", "lci", "diff", 
    "t-stat"), value = format(c(x.N_star, sd.N_star, se.N_star, lci.N_star, 
    uci.N_star, diff.N_star, t.N_star), scientific = FALSE, digits = 2))
N.table
```

Note: The original lab called for histograms, however, histograms can be strongly affected by the number of bins used, so I've opted for density plots.
```{r, message=FALSE}
df <- data.frame(m2=m2) # create a dataframe with m2values
x <- round(seq(5,35, length.out=length(df$m2)))
df$m2n <- rnorm(x, mean = Em2, sd = sd.m2)# create a normal draw with mean of E(m2) and sd of m2

df <- melt(df)

ggplot(df, aes(x=value, group=variable, fill=variable))+ geom_density(alpha=.3)+ 
                        scale_fill_manual(name="",
                         breaks=c("m2", "m2n"),
                         labels=c("Observed", "Predicted"),
                         values=c('black', 'blue'))+xlab('M2')



df <- data.frame(N_star=N_star) # create a dataframe with N* values
x <- seq(30000,230000, length.out=length(df$N_star)) 
df$N_starn <- rnorm(x, mean = N, sd = sd.N_star)# create a normal draw with mean of N and sd of N*
df <- melt(df)

 ggplot(df, aes(x=value, group=variable, fill=variable))+ geom_density(alpha=.3)+ 
                        scale_fill_manual(name="",
                         breaks=c("N_star", "N_starn"),
                         labels=c("Observed", "Predicted"),
                         values=c('black', 'blue'))+xlab('N*')
```

Notes: A reasonably normal distribution of draws of $M_2$ produce a skewed distribution of population estimates.

```{r}
sessionInfo()
```


This is a lab exercise from Dr. Terry Quinn  
at the University of Alaska Fairbanks  
and has been worked up in R by:  
Ben Williams  
bcwilliams2@alaska.edu  
ben.williams@alaska.gov
